I will use the Machine Learning Workflow to process and transform DOT data to create a prediction mode. This model must predict whether a filght would arrive 15+ minutes after the scheduled arrival time with 70% accuracy.
```{r}
FlightData <- read.csv('196248054_T_ONTIME_REPORTING.csv', sep=",", header = T, stringsAsFactors = F)
nrow(FlightData)
```
Too many records so I will take subset from that
```{r}
airports <- c('ATL', 'LAX', 'ORD', 'DFW', 'JFK','SFO','CLT','LAS','PHX')
FlightData <- subset(FlightData, DEST %in% airports & ORIGIN %in% airports)
nrow(FlightData)
head(FlightData,2)
```
As you can see the last column is usless so I will remove it.
```{r}
FlightData$X <- NULL
```
Now I am looking for correlated kolumns to remove kolumns.
```{r}
cor(FlightData[c("ORIGIN_AIRPORT_SEQ_ID", "ORIGIN_AIRPORT_ID")])
cor(FlightData[c("DEST_AIRPORT_SEQ_ID", "DEST_AIRPORT_ID")])
FlightData$ORIGIN_AIRPORT_ID <- NULL
FlightData$DEST_AIRPORT_ID <- NULL
```
Carrier also need to be removed because they are the same.
```{r}
FlightData$OP_UNIQUE_CARRIER <- NULL
```
Now I will remove rows which don't contain 1 lub 0 in Arr_Del15, Dep_Del15 column.

```{r}
FlightDataUpdate <- FlightData[!is.na(FlightData$ARR_DEL15) & FlightData$ARR_DEL15 != "" & !is.na(FlightData$DEP_DEL15) & FlightData$DEP_DEL15 != "",]
nrow(FlightDataUpdate)
```
In this step I will convert values from string to int or factor
```{r}
FlightDataUpdate$DISTANCE <- as.integer(FlightDataUpdate$DISTANCE)
FlightDataUpdate$CANCELLED <- as.integer(FlightDataUpdate$CANCELLED)
FlightDataUpdate$DIVERTED <- as.integer(FlightDataUpdate$DIVERTED)
FlightDataUpdate$ARR_DEL15 <- as.factor(FlightDataUpdate$ARR_DEL15)
FlightDataUpdate$DEP_DEL15 <- as.factor(FlightDataUpdate$DEP_DEL15)
FlightDataUpdate$DEST_AIRPORT_SEQ_ID <- as.factor(FlightDataUpdate$DEST_AIRPORT_SEQ_ID)
FlightDataUpdate$ORIGIN_AIRPORT_SEQ_ID <- as.factor(FlightDataUpdate$ORIGIN_AIRPORT_SEQ_ID)
FlightDataUpdate$DAY_OF_WEEK <- as.factor(FlightDataUpdate$DAY_OF_WEEK)
FlightDataUpdate$DEST <- as.factor(FlightDataUpdate$DEST)
FlightDataUpdate$ORIGIN <- as.factor(FlightDataUpdate$ORIGIN)
FlightDataUpdate$DEP_TIME_BLK <- as.factor(FlightDataUpdate$DEP_TIME_BLK)
FlightDataUpdate$OP_CARRIER <- as.factor(FlightDataUpdate$OP_CARRIER)
```
Look how many rows contain 1 (true) or 0 (false)
```{r}
tapply(FlightDataUpdate$ARR_DEL15, FlightDataUpdate$ARR_DEL15, length)
```
Now I select 6 columnts to build model machine learning: 
-Origin
-Destination
-Day of week
-Carrier
-Departure time block
-Arrival delay 15
```{r}
set.seed(133515)
ImportantColumn <- c("ARR_DEL15", "DAY_OF_WEEK","OP_CARRIER","DEST","ORIGIN","DEP_TIME_BLK")
NewFlightDataUpdate <- FlightDataUpdate[,ImportantColumn]
Trainrow <- createDataPartition(NewFlightDataUpdate$ARR_DEL15, p=0.70, list = F)

```
Start training data
```{r}
library(e1071)
TrainFlightDataUpdate <- NewFlightDataUpdate[Trainrow,]
TestData <- NewFlightDataUpdate[-Trainrow,]
```
I will use random forest now:
```{r}
randomforestModel <- randomForest(TrainFlightDataUpdate[-1], TrainFlightDataUpdate$ARR_DEL15, proximity = T, importance = T)
randomforestValidation <- predict(randomforestModel, TestData)
randomforestConfusionMatrix <- confusionMatrix(randomforestValidation, TestData[,"ARR_DEL15"])
randomforestConfusionMatrix


```


